---
title: "Azimuth and elevation model creation"
author: "Richard Massy"
date: "2023-07-21"
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

1.0 Figure 1 - Virtual flight paths: load functions

File dependencies: Sun position detail.csv
Environment dependencies: None

Functions:

pathPrep() loads the azimuths of the days/ location combo into a dataframe which is saved to the environment as "s" and saves an empty variable "Tracks" to the environment which will contain the subsequent migration tracks

compPath() simulates migration tracks for the supplied compensation levels and saves them to "tracks"

compInaccPath() simulates migration tracks for the supplied compensation levels with added inaccuracy and saves them to "tracks"

avgPath(), avgInaccPath(), jumpPath() and jumpInaccPath() do the same for time averaging and step compensation respectively

plotTrack() takes the "tracks" global variable and visualises it, producing figure 1a
```{r}
pathPrep = function(days=c("180901","180902","180903","180904","180905"),
                    location="Bujaruelo"){
  sunPosDat = subset(read.csv("Sun position detail.csv"),Location==location)
  s <<- do.call(rbind,lapply(days,function(day){
    hours = sunPosDat[sunPosDat$Date==day,]$Time
    s1 = data.frame("minutes"=seq(min(hours),max(hours),1/1440))
    model1 = smooth.spline(sunPosDat[sunPosDat$Date==day,c("Time","Azimuth")],
                           all.knots=T)
    s1$azimuths = predict(model1,s1$minutes)$y/180*pi
    return(s1)
  }))
  tracks <<- NULL
}
# Fullcomp
compPath = function(comp_levels = c(0,0.7,1)){
  for (level in comp_levels){
    comp_vectors = (s$azimuths-level*(s$azimuths-pi))%%(2*pi)
    track = matrix(rep(c(0,rep(NA,nrow(s))),2),
                   ncol=2,dimnames=list(NULL,c("x","y")))
    for (i in seq_along(comp_vectors)){
      track[i+1,] = c(track[i,"x"] + sin(comp_vectors[[i]]),
                      track[i,"y"] + cos(comp_vectors[[i]]))
    }
    tracks[[as.character(level)]] <<- list(track)
  }
}
# Fullcomp + inacc
compInaccPath = function(comp_levels=c(0,0.5,1),nTrack=5,kapp=3.289){
  if (!"circular" %in%(.packages())){require(circular)}
  for (level in comp_levels){
    track = list()
    for (n in 1:nTrack){
      inacc=rvonmises(nrow(s),circular(0,unit="radians",template="geographics",
                                       modulo="2pi"),kappa = kapp)
      comp_vectors = (s$azimuths-level*(s$azimuths-pi)+inacc)%%(2*pi)
      trackN = matrix(rep(c(0,rep(NA,nrow(s))),2),
                      ncol=2,dimnames=list(NULL,c("x","y")))
      for (i in seq_along(comp_vectors)){
        trackN[i+1,] = c(trackN[i,"x"] + sin(comp_vectors[[i]]),
                         trackN[i,"y"] + cos(comp_vectors[[i]]))
      }
      track[[paste(n)]] = trackN
    }
    tracks[[paste0(level,"*")]] <<- track
  }
}
# Time averaging (optimum (no inacc) 1-5 Sep '18 = 19.375)
avgPath = function(step_levels = c(19.1*0.7,19.1),zenith = 0.5829041){
  for (level in step_levels){
    step_vectors = (s$azimuths+(zenith-s$minutes)*level/7.5*pi)%%(2*pi)
    track = matrix(rep(c(0,rep(NA,nrow(s))),2),
                   ncol=2,dimnames=list(NULL,c("x","y")))
    for (i in seq_along(step_vectors)){
      track[i+1,] = c(track[i,"x"] + sin(step_vectors[[i]]),
                      track[i,"y"] + cos(step_vectors[[i]]))
    }
    tracks[[paste0(level,"째/h")]] <<- list(track)
  }
}
# Time averaging + Inacc
avgInaccPath = function(step_levels=c(19.1),nTrack=5,zenith=0.5829041,kapp=3.289){
  if (!"circular" %in%(.packages())){require(circular)}
  for (level in step_levels){
    track = list()
    for (n in 1:nTrack){
      inacc=rvonmises(nrow(s),circular(0,unit="radians",template="geographics",
                                       modulo="2pi"),kappa = kapp)
      step_vectors = (s$azimuths+(zenith-s$minutes)*level/7.5*pi+inacc)%%(2*pi)
      trackN = matrix(rep(c(0,rep(NA,nrow(s))),2),
                      ncol=2,dimnames=list(NULL,c("x","y")))
      for (i in seq_along(step_vectors)){
        trackN[i+1,] = c(trackN[i,"x"] + sin(step_vectors[[i]]),
                         trackN[i,"y"] + cos(step_vectors[[i]]))
      }
      track[[paste(n)]] = trackN
    }
    tracks[[paste0(level,"째/h*")]] <<- track
  }
}


# Jumpcomp (optimum (no inacc) 1-5 Sep '18 = 121.2)
jumpPath = function(jump_levels = c(121*0.7,121),zenith = 0.5829041){
  for (level in jump_levels){
    lvRad = level/180*pi
    jump_vectors = (s$azimuths + round(0.5+zenith-s$minutes)*lvRad - lvRad/2)%%(2*pi)
    track = matrix(rep(c(0,rep(NA,nrow(s))),2),
                   ncol=2,dimnames=list(NULL,c("x","y")))
    for (i in seq_along(jump_vectors)){
      track[i+1,] = c(track[i,"x"] + sin(jump_vectors[[i]]),
                      track[i,"y"] + cos(jump_vectors[[i]]))
    }
    tracks[[paste0(level,"째")]] <<- list(track)
  }
}

# Jumpcomp + Inacc
jumpInaccPath = function(jump_levels=c(121),nTrack=5,zenith=0.5829041,kapp=3.289){
  if (!"circular" %in%(.packages())){require(circular)}
  for (level in jump_levels){
    track = list()
    lvRad = level/180*pi
    for (n in 1:nTrack){
      inacc=rvonmises(nrow(s),circular(0,unit="radians",template="geographics",
                                       modulo="2pi"),kappa = kapp)
      jump_vectors =(s$azimuths+round(0.5+zenith-s$minutes)*lvRad-lvRad/2+inacc)%%(2*pi)
      trackN = matrix(rep(c(0,rep(NA,nrow(s))),2),
                      ncol=2,dimnames=list(NULL,c("x","y")))
      for (i in seq_along(step_vectors)){
        trackN[i+1,] = c(trackN[i,"x"] + sin(jump_vectors[[i]]),
                         trackN[i,"y"] + cos(jump_vectors[[i]]))
      }
      track[[paste(n)]] = trackN
    }
    tracks[[paste0(level,"째*")]] <<- track
  }
}

# Plotting virtual flight path
plotTrack = function(b0=0.14,b1=0.7,b2=2){
  par(pty="s",mar=c(4.1,4.1,3.1,2.1)) # Mar order = bottom, left, top, right
  plot(NULL, xlim=c(0,length(tracks)*0.2-0.15),ylim=c(1,0),xaxt="n",las=1,asp=1,
       ylab="Southward efficiency",xlab="Compensation level",
       main="Simulated flight paths 1-5 September 2018")
  polygon(x=c(b0,b1,b1,b0),y=c(-1,-1,2,2),border=NA,
          col=rgb(0,255,50,alpha=30,max=255))
  polygon(x=c(b1,b2,b2,b1),y=c(-1,-1,2,2),border=NA,
          col=rgb(255,250,0,alpha=50,max=255))
  if (!b2==2){
    polygon(x=c(b2,2,2,b2),y=c(-1,-1,2,2),border=NA,
            col=rgb(0,0,250,alpha=30,max=255))
  }
  abline(h=c(0,1),lty=2,col="grey")
  for (i in seq_along(tracks)){
    for (j in seq_along(tracks[[i]])){
      tracks[[i]][[j]] = tracks[[i]][[j]]/3925
      tracks[[i]][[j]][,"x"] = tracks[[i]][[j]][,"x"] + (i-1)/5
      lines(-y~x,data=tracks[[i]][[j]],lwd=0.6)
    }
  }
  axis(1,at=seq(0,(i-1)/5,length.out=length(tracks)),labels=names(tracks))
}

```
1.1a Figure 1a - Virtual flight paths running: visualisation of full, time averaging and step compensation

File dependencies: Sun position detail.csv(downloaded)
Environment dependencies: pathPrep + compPath + avgPath + jumpPath + plotTrack(all from 1.0)
```{r}

pathPrep()
compPath()
avgPath()
jumpPath()
plotTrack(b0=0.13,b1=0.5,b2=0.9)
```

1.1b Virtual flight paths running: old version including inaccuracy visualisation

File dependencies: Sun position detail.csv(downloaded)
Environment dependencies: pathPrep + compPath + compInaccPath + avgPath + avgInaccPath + plotTrack(all from 1.0)
```{r}
pathPrep()
compPath()
compInaccPath(comp_levels = 1, nTrack = 3)
avgPath()
avgInaccPath(step_levels = 19.4, nTrack = 3)
plotTrack()
```

1.2 visualising the azimuth change over time for different locations showing the how the real and estimated azimuths compare for the different compensation strategies

File dependencies: Sun position detail.csv(downloaded)
Environment dependencies: pathPrep(1.0)

1.2a load plotComp() function and run an example (time averaging, Bujaruelo)
```{r}
plotComp = function(df=s,zenith=0.5835,CL=1,AL=17.6,SL=123,methods="full",partial=F,
                    colour=F){
  df$hours = df$minutes*24
  df$azimuths = df$azimuths*180/pi
  equations= list("full"=function(p=1){180+CL*p*(df$azimuths-180)},
                  "avg"=function(p=1){180-((zenith-df$minutes)*AL*p*24)},
                  "step"=function(p=1){180-(round(0.5+zenith-df$minutes)*SL*p-SL*p/2)})
  colours = list("full"=rgb(0,255,50,alpha=30,max=255),
              "avg"=rgb(255,250,0,alpha=50,max=255),
              "step"=rgb(0,0,250,alpha=30,max=255))
  if (colour){
    plot(azimuths~hours,data=df,type="n",
         xaxt="n",yaxt="n",xlab="",ylab="")
    rect(par("usr")[1], par("usr")[3],
         par("usr")[2], par("usr")[4],
         col = colours[[methods[1]]])
    par(new=T)
  }
  plot(azimuths~hours,data=df,type="l",lwd=2,col="red",
       xlab="Time (hours)",ylab="Azimuth")
  abline(h=180,lty=2,col="grey")
  for (method in methods){
    df$vectors = equations[[method]]()
    lines(vectors~hours,data=df,type="l",lwd=2,col="blue",lty="dashed")
    if (partial){
      df$vectors = equations[[method]](p=partial)
      lines(vectors~hours,data=df,type="l",lwd=2,col="deepskyblue",lty="dashed")
    }
  }
}
pathPrep(days="180901")
plotComp(method="avg")
```

1.2b Figure 1b - run the azimuth visualisation for Bujaruelo and Panama

NOTE: graph dimensions are designed for export so appear stretched

```{r}
pathPrep(days="180901")
par(mfrow=c(3,1),mar=c(5-1,4,4-3,2-1)+0.1,cex.axis=1.3,cex.lab=1.4)
lapply(c("full","avg","step"),function(method){
  plotComp(methods=method,partial=0.7,colour=T)})


pathPrep(days="180901",location="Neotropical")
lapply(c("full","avg","step"),function(method){
  plotComp(zenith=0.555 ,AL=22.8,SL=180,methods=method,partial=0.7)})
```

1.2b Figure 1b - export the azimuth visualisations for Falsterbo, Bujaruelo, Maghreb and Panama in proper dimensions

Saves:
"RGraphCompDemoFal" (file of the azimuth visualisations for Falsterbo)
"RGraphCompDemoBuj" (file of the azimuth visualisations for Bujaruelo)
"RGraphCompDemoMag" (file of the azimuth visualisations for Maghreb)
"RGraphCompDemoNeo" (file of the azimuth visualisations for Panama)

NOTE: takes a few seconds to run & might have compatability issues saving images

SECOND NOTE: if adjusting the code, the zenith time, adjustment rate (AL) and step size (SL) must be manually input into the plotComp() function. The default is Bujaruelo.

Zeniths: Falsterbo 0.5476, Bujaruelo 0.5835, Maghreb 0.5217, Subtropical 0.5553
AL: Falsterbo 16.4, Bujaruelo 17.6, Maghreb 19.0, Subtropical 22.8
SL: Falsterbo 118, Bujaruelo 123, Maghreb 134, Subtropical 180
```{r}
pathPrep(days="180901",location="Falsterbo")
png(file="RGraphCompDemo1Fal.png",res=1000,width=3200,height=7200,pointsize=12,
    antialias="cleartype")
par(mfrow=c(3,1),mar=c(5-1,4,4-3,2-1)+0.1,cex.axis=1.4,cex.lab=1.8)
lapply(c("full","avg","step"),function(method){
  plotComp(zenith=0.548 ,AL=16.4,SL=118,methods=method,partial=0.7)})
dev.off()

pathPrep(days="180901")
png(file="RGraphCompDemo2Buj.png",res=1000,width=3200,height=7200,pointsize=12,
    antialias="cleartype")
par(mfrow=c(3,1),mar=c(5-1,4,4-3,2-1)+0.1,cex.axis=1.4,cex.lab=1.8)
lapply(c("full","avg","step"),function(method){
  plotComp(zenith=0.584 ,AL=17.6,SL=123,methods=method,partial=0.7,colour=T)})
dev.off()

pathPrep(days="180901",location="Maghreb")
png(file="RGraphCompDemo3Mag.png",res=1000,width=3200,height=7200,pointsize=12,
    antialias="cleartype")
par(mfrow=c(3,1),mar=c(5-1,4,4-3,2-1)+0.1,cex.axis=1.4,cex.lab=1.8)
lapply(c("full","avg","step"),function(method){
  plotComp(zenith=0.522 ,AL=19.0,SL=134,methods=method,partial=0.7)})
dev.off()

pathPrep(days="180901",location="Neotropical")
png(file="RGraphCompDemo4Neo.png",res=1000,width=3200,height=7200,pointsize=12,
    antialias="cleartype")
par(mfrow=c(3,1),mar=c(5-1,4,4-3,2-1)+0.1,cex.axis=1.4,cex.lab=1.8)
lapply(c("full","avg","step"),function(method){
  plotComp(zenith=0.555 ,AL=22.8,SL=180,methods=method,partial=0.7)})
dev.off()
```